---
title: "medstats: Quick Start Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{medstats: Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

**medstats** is a comprehensive R package for medical statistics, providing streamlined functions for:

- **Table formatting**: Publication-ready three-line tables and Word export
- **Data processing**: Survival data conversion, baseline tables
- **Regression analysis**: Automated GLM and Cox regression with univariate/multivariate results
- **Repeated measures**: GEE analysis for longitudinal data
- **Publication-quality plots**: KM curves, forest plots, RCS splines, ROC curves, Sankey diagrams, and more

```{r setup}
library(medstats)
```

# Table Formatting

## Format a flextable

The `format_flextable()` function applies a unified three-line table style:

```{r eval=FALSE}
# From a data frame
ft <- format_flextable(head(mtcars[, 1:5]))

# From a gtsummary object
library(gtsummary)
tbl <- tbl_summary(trial, include = c(age, grade, trt), by = trt)
ft <- format_flextable(tbl)
```

## Export tables to Word

Use `export_word()` to batch-export multiple tables into a single `.docx` file:

```{r eval=FALSE}
export_word(
  data_list = list(head(mtcars), head(iris)),
  table_titles = c("Table 1: mtcars", "Table 2: iris"),
  output_file = "my_tables.docx"
)
```

# Data Processing

## Convert longitudinal data to survival format

```{r eval=FALSE}
surv_data <- long_to_surv_data(
  data = my_clinical_data,
  id_var = "PatientID",
  event_flag_var = "is_event",
  time_var = "follow_up_time",
  baseline_vars = c("Sex", "Age")
)
```

## Create a baseline table (Table 1)

```{r eval=FALSE}
make_table1(
  data = trial,
  vars = c("age", "marker", "stage", "grade"),
  specific_vars = c("marker"),  # Non-normal distribution
  group_var = "trt"
)
```

# Regression Analysis

## Automated GLM regression

```{r eval=FALSE}
# Linear regression
run_glm_auto(mtcars, vars = c("hp", "wt"), outcome_var = "mpg", family = "gaussian")

# Logistic regression
run_glm_auto(trial, vars = c("age", "stage"), outcome_var = "response", family = "binomial")
```

## Automated Cox regression

```{r eval=FALSE}
run_cox_auto(lung, vars = c("age", "sex"), time_var = "time", event_var = "status")
```

# Plotting

## Mean ± SE line chart

```{r eval=FALSE}
plot_meanse(
  data = ChickWeight,
  target_var = "weight",
  time_var = "Time",
  group_var = "Diet"
)
```

## KM survival curve

```{r eval=FALSE}
lung$status2 <- ifelse(lung$status == 2, 1, 0)
plot_km(lung, group_var = "sex", time_var = "time", status_var = "status2")
```

## RCS restricted cubic spline

```{r eval=FALSE}
plot_rcs(mtcars, exposure = "wt", outcome = "mpg", model_type = "linear")
```

## ROC curve

```{r eval=FALSE}
model <- glm(am ~ mpg + hp + wt, data = mtcars, family = binomial)
mtcars$pred <- predict(model, type = "response")
plot_roc(mtcars, true_var = "am", pred_var = "pred")
```
